Gammapy: A Python package for gamma-ray astronomy

### 2022-11-30 src/ms-review.tex
Address more comments by C.Nigro

\subsection{Broadband SED Modeling}
\label{ssec:broadband-sed-modeling}
By combining \gammapy with astrophysical modelling codes, users can also fit
astrophysical spectral models to \gammaray data. In \gammaray
astronomy one typically observes two radiation production
mechanisms, the so-called hadronic and leptonic scenarios.
There are mutiple Python packages that are able to model
the \gammaray emission, given a physical scenario. Among those
packages are  Agnpy \citep{agnpy}, Naima \citep{naima}, Jetset \citep{jetset}
and Gamera \citep{gamera}.
Tyically those emission models predict broadband emission from
radio, up to the very high-energy \gammaray range.
By relying on the multiple dataset types in \gammapy those
data can be combined to constrain such a broadband emission model.
\gammapy provides a built-in \code{NaimaSpectralModel} that allows
users to wrap a given astrophysical emission model from the
naima package and fit it directly to \gammaray data.

@@

In this section, we provide an overview of the organization of the \gammapy
project. We briefly describe the main roles and responsibilities within the
team, as well as the technical infrastructure designed to facilitate the
development and maintenance of \gammapy as a high quality software. We use
common tools and services for software development of Python open-source
projects, code review, testing, package distribution and user support, with a
customized solution for a versioned thoroughly tested documentation in the form
of user-friendly playable tutorials. This section concludes with an outlook on
the roadmap for future directions.

@@

\gammapy is a well-organised international open-source project with a broad
developer base and the presence of commitments from strong groups and
institutes in the very high-energy astrophysics
domain\footnote{\url{https://gammapy.org/team.html}}. The main development
roadmaps are discussed and validated by a Coordination Committee, composed of
representatives of the main contributing laboratories. This committee is
chaired by a Project Manager and his deputy while two Lead Developers manage
the development strategy and organise technical activities. Because of this
institutionally driven characteristic, the permanent staff and commitment of
supporting institutes ensure the continuity of the executive teams. A core team
of developers from the contributing laboratories is in charge of the regular
development, which benefits from punctual contributions of the community at
large.

@@

Several automated tasks are set as \github
actions\footnote{\url{https://\github.com/features/actions}}, blocking the
processes and alerting developers when fails occur. This is the case of the
content integration workflow, which monitors the execution of the test coverage
suite\footnote{\url{https://pytest.org}} using datasets from the
\textit{gammapy-data} repository. Tests scan not only the codebase but also the
code snippets present in docstrings of the scripts and in the RST documentation
files, as well as in the tutorials provided in the form of Jupyter notebooks.

@@

\subsection{Release Cycle, Versioning, and Long-term Support}
\label{ssec:release-cycle}
With the first long term support (LTS) release v1.0, the \gammapy project
enters a new development phase. The development will change from
quick feature driven development to more stable maintenance
and user support driven developement. After v1.0 we foresee
a developement cycle with major, minor and bugfix releases.
We basically followg the development cycle of the Astropy
project. Thus we expect a major LTS release approximately
every two years, minor releases are planned every 6~months,
while bug-fix releases will happen as needed. While
bug-fix releases will not introduce API breaking changes,
we will work with a deprecation system for minor releases.
API breaking changes will be announced to user by runtime
warnings first and then implemented in the subsequent
minor release. This approach we consider a fair
compromise between the interests of users in a stable
package and the interest of developers to improve
and develop \gammapy in future. The developemnt cycle is described
in more detail in PIG 24 \todo{reference}.

\section{Reproducibility}
\label{sec:reproducibility}
One of the most important goals of the \gammapy project is to support open and
reproducible science results. Thus we decided to write this manuscript
openly and publish the Latex source code along with the associated
Python scripts to create the figures
in \url{https://\github.com/gammapy/gammapy-v1.0-paper}.
This \github repository also documents the history of the creation
and evolution of the manuscript with time. To simplify the reproducibility
of this manuscript including figures and text, we relied on the tool
"showyourwork"~\citep{Luger2021}. This tool coordinates the building
process and software as well as data dependencies such, that the complete
manuscript can be reproduced with a single \code{make} command, after
downloading the source repository. See \todo{reference?} for detailed
instructions. Almost all figures in this manuscript provide a link
to a Python script, that was used to produce it. This means all
example analyses presented in Sec.\ref{sec:applications} link to
actually working Python source code.

\section{Summary and Outlook}
\label{sec:summary-and-outlook}
%
In this manuscript we presented the first LTS version of \gammapy.
\gammapy is a Python package for \gammaray astronomy, which relies on the
scientific Python ecosystem, including Numpy and Scipy and Astropy as
main dependencies. It also holds the status of an Astropy affiliated
package. It supports high-level analysis of astronomical \gammaray
data from intermediate level data formats, such as the FITS based
\gadf. Starting from lists of \gammaray events and corresponding description
of the instrument response users can reduce and project the data
to WCS, HEALPix and region based data structures. The reduced data is bundled
into datasets, which serve as a basis for Poisson maximum likelihood
modelling of the data. For this purpose \gammapy provides a wide selection
of built-in, spectral, spatial and temporal models as well as unified
fitting interface with connection to multiple optimization backends.

>>>>

\subsection{Broadband SED Modeling}
\label{ssec:broadband-sed-modeling}
By combining \gammapy with astrophysical modelling codes, users can also fit
astrophysical spectral models to \gammaray data. In \gammaray
astronomy one typically observes two radiation production
mechanisms, the so-called hadronic and leptonic scenarios.
There are several Python packages that are able to model
the \gammaray emission, given a physical scenario. Among those
packages are  Agnpy \citep{agnpy}, Naima \citep{naima}, Jetset \citep{jetset}
and Gamera \citep{gamera}.
Tyically those emission models predict broadband emission from
radio, up to the very high-energy \gammaray range.
By relying on the multiple dataset types in \gammapy those
data can be combined to constrain such a broadband emission model.
\gammapy provides a built-in \code{NaimaSpectralModel} that allows
users to wrap a given astrophysical emission model from the
naima package and fit it directly to \gammaray data.

As an example of this application, we use the same multi-instrument
dataset described in the previous section and we fit it with an inverse
Compton model computed with naima and wrapped in the \gammapy fitting
 routine with the NaimaSpectraModel class. We consider a log-parabolic
electron distribution scattering: the synchrotron radiation produced by
the same electrons; near and far infrared photon fields and the cosmic
microwave background (CMB). We adopted the prescription on the target
photon fields provided in the naima 
documentation \footnote{\url{https://naima.readthedocs.io/en/latest/examples.html#crab-nebula-ssc-model}}.
The best-fit inverse Compton spectrum is represented with a red dashed line in
Fig.~\ref{fig:multi_instrument_analysis}.

@@

In this section, we provide an overview of the organization of the \gammapy
project. We briefly describe the main roles and responsibilities within the
team, as well as the technical infrastructure designed to facilitate the
development and maintenance of \gammapy as a high-quality software. We use
common tools and services for software development of Python open-source
projects, code review, testing, package distribution and user support, with a
customized solution for a versioned and thoroughly-tested documentation in the form
of user-friendly playable tutorials. This section concludes with an outlook on
the roadmap for future directions.

@@

\gammapy is an international open-source project with a broad
developer base and contributions and commitments from mutiple groups and
leading institutes in the very high-energy astrophysics
domain\footnote{\url{https://gammapy.org/team.html}}. The main development
roadmaps are discussed and validated by a \textit{Coordination Committee}, composed of
representatives of the main contributing institutions and observatories.
This committee is
chaired by a \textit{Project Manager} and his deputy while two \textit{Lead Developers} manage
the development strategy and organise technical activities. This
institutionally driven organisation, the permanent staff and commitment of
supporting institutes, ensures the continuity of the executive teams. A core team
of developers from the contributing institutions is in charge of the regular
development, which benefits from punctual contributions of the community at
large.

@@

Several automated tasks are set as \github
actions\footnote{\url{https://\github.com/features/actions}}, blocking the
processes and alerting developers when fails occur. This is the case of the
continuous integration workflow, which monitors the execution of the test coverage
suite\footnote{\url{https://pytest.org}} using datasets from the
\textit{gammapy-data} repository. Tests scan not only the codebase but also the
code snippets present in docstrings of the scripts and in the RST documentation
files, as well as in the tutorials provided in the form of Jupyter notebooks.

@@

\subsection{Release Cycle, Versioning, and Long-term Support}
\label{ssec:release-cycle}
With the first long term support (LTS) release v1.0, the \gammapy project
enters a new development phase. The development will change from
quick feature-driven development to more stable maintenance
and user support driven developement. After v1.0 we foresee
a developement cycle with major, minor and bugfix releases;
we basically follow the development cycle of the Astropy
project. Thus we expect a major LTS release approximately
every two years, minor releases are planned every 6~months,
while bug-fix releases will happen as needed. While
bug-fix releases will not introduce API breaking changes,
we will work with a deprecation system for minor releases.
API-breaking changes will be announced to user by runtime
warnings first and then implemented in the subsequent
minor release. This approach we consider a fair
compromise between the interests of users in a stable
package and the interest of developers to improve
and develop \gammapy in future. The development cycle is described
in more detail in PIG 23 \citep{gammapy_pig_23}.

\section{Reproducibility}
\label{sec:reproducibility}
One of the most important goals of the \gammapy project is to support open and
reproducible science results. Thus we decided to write this manuscript
openly and publish the Latex source code along with the associated
Python scripts to create the figures
in \url{https://github.com/gammapy/gammapy-v1.0-paper}.
This \github repository also documents the history of the creation
and evolution of the manuscript with time. To simplify the reproducibility
of this manuscript including figures and text, we relied on the tool
\textit{showyourwork}~\citep{Luger2021}. This tool coordinates the building
process and software as well as data dependencies such, that the complete
manuscript can be reproduced with a single \code{make} command, after
downloading the source repository. See \url{https://github.com/gammapy/gammapy-v1.0-paper/blob/main/README.md} for detailed
instructions. Almost all figures in this manuscript provide a link
to a Python script, that was used to produce it. This means all
example analyses presented in Sec.\ref{sec:applications} link to
actually working Python source code.

\section{Summary and Outlook}
\label{sec:summary-and-outlook}
%
In this manuscript we presented the first LTS version of \gammapy.
\gammapy is a Python package for \gammaray astronomy, which relies on the
scientific Python ecosystem, including Numpy and Scipy and Astropy as
main dependencies. It also holds the status of an Astropy affiliated
package. It supports high-level analysis of astronomical \gammaray
data from intermediate level data formats, such as the FITS based
\gadf. Starting from lists of \gammaray events and corresponding description
of the instrument response users can reduce and project the data
to WCS, HEALPix and region based data structures. The reduced data is bundled
into datasets, which serve as a basis for Poisson maximum likelihood
modelling of the data. For this purpose \gammapy provides a wide selection
of built-in spectral, spatial and temporal models, as well as unified
fitting interface with connection to multiple optimization backends.
